Detector-frame cosmology with non-parametric methods

### 2024-04-12 src/tex/ms.tex
Add method framework

\begin{acronym}
    \acro{GW}{gravitational wave}
    \acro{LVK}{LIGO-Virgo-KAGRA Collaboration}
    \acro{PE}{parameter estimation}
    \acro{CMB}{cosmic microwave background}
    \acro{EM}{electromagnetic}
\end{acronym}

@@

@@

@@

@@

@@

>>>>

\begin{acronym}
    \acro{GW}{gravitational wave}
    \acro{LVK}{LIGO-Virgo-KAGRA Collaboration}
    \acro{PE}{parameter estimation}
    \acro{CMB}{cosmic microwave background}
    \acro{EM}{electromagnetic}
    \acro{DPGMM}{Dirichlet process Gaussian mixture model}
    \acro{BBH}{binary black hole}
    \acro{GWTC-3}{the third Gravitational-Wave Transient Catalog}
\end{acronym}

@@

The first step of our method is to reconstruct the observed population.
we choose to consider the primary mass $m_1$ of the \ac{BBH} sources.
From the posterior samples obtained from the \ac{PE} of each \ac{GW} event, we can reconstruct the observed detector-frame mass distribution $p(m_1|\{d\})$, where $\{d\}$ is the set of all the observed data.

To perform the reconstruction, we use a non-parametric reconstruction package \textsc{FIGARO} \citep{Rinaldi:2022kyg}.
\textsc{FIGARO} can reconstruct an arbitrary probability density function from a set of samples based on the \ac{DPGMM}.
It can also perform a hierarchy of \ac{DPGMM} to reconstruct an polulation distribution with posterior samples from multiple events \citep{Rinaldi:2021bhm}.
The output of \textsc{FIGARO} is a set of distributions, where each distribution is a realization of the observed population that is consistent with the data.
Therefore, the set of distributions represents the uncertainty in the reconstruction procedure.
In Fig.~XXX \todo{Add figure}, we show an example of the reconstruction obtained with \textsc{FIGARO}.

This part of the method is completely data-driven and does not require any assumptions about the population model or the cosmological model.
Therefore the intermediate result is independent of the assumptions, and we can reuse the result for different population models, cosmological models, or selection functions.
This reduces the computational cost of the method, as we only need to perform the reconstruction of the observed population once.

@@

On the other hand, we need a source-frame population model $p(m_1|\Lambda)$ to represent the underlying population of the sources, where $\Lambda$ is the polulation model.
We can either use a parametric model to only assume the form of the population distribution, or we can use a fixed population model to assume the exact population distribution.
With $p(m_1|\Lambda)$, we need to transform it to the detector frame with a given cosmological model to obtain the detector-frame population model $p(m_1|\Lambda, \Omega)$, where $\Omega$ is the cosmological model.
In this study, we use the $\Lambda$CDM model with the Planck 2018 cosmological parameters \citep{Planck:2018vyg}, except for $H_0$ which we choose to be a free parameter.
With different $\Omega$, $p(m_1|\Lambda)$ will be transformed differently, and the resulting $p(m_1|\Lambda, \Omega)$ will be different.
In Fig.~XXX \todo{Add figure}, we show an example of the transformation without any selection effects.

After the cosmological transformation, we need to consider the selection effects.
The selection effects can be easily incorporated by applying the selection function to the transformed population.
\todo{Add more details. Parametric or fixed?}

Opposite to the first part of the method, which is completely data-driven, this part includes all the assumptions about the population model, the cosmological model, and the selection function.

@@

With $p(m_1|\{d\})$ and $p(m_1|\Lambda, \Omega)$ obtained from Sec.~\ref{sec:reconstruction} and Sec.~\ref{sec:transformation} respectively, we can perform the \ac{PE} to infer the free parameters in the parametric models assumed in Sec.~\ref{sec:transformation}.
The \ac{PE} is performed by comparing the observed population $p(m_1|\{d\})$ with the transformed population $p(m_1|\Lambda, \Omega)$.
To compare the two populations, we use the Jensen–Shannon distance $d_{JS}$ as the distance measure.
$d_{JS}$ is calculated as
\begin{equation}
    d_{JS}(p, q) = \sqrt{\frac{D_{KL}(p||m) + D_{KL}(q||m)}{2}},
\end{equation}
where $p$ and $q$ are the two distributions to be compared, $m = (p + q) / 2$, and $D_{KL}$ is the Kullback–Leibler divergence.
$d_{JS}$ is a symmetric and smoothed version of $D_{KL}$, and it is bounded between 0 and 1.
For each reconstructed population $p(m_1|\{d\})$, we find the best-fit parameters by minimizing $d_{JS}$ with respect to the free parameters.
This procedure gives us a set of best-fit parameters for each reconstructed population.
Note that the resulting distribution of the best-fit parameters is not a posterior distribution in the Bayesian sense, as we do not consider the likelihood of the observed data given the population model.
Instead, it is a distribution of the best-fit parameters that are consistent with the observed data, distributed according to the uncertainty in the reconstruction of the observed population.
In Fig.~XXX \todo{Add figure}, we show an example of the fitting procedure.

@@

\subsection{GWTC-3 test}
\label{sec:gwtc3}

@@

% Comparison with other methods

% Future directions
    % Likelihood
    % luminosity distance/redshift polulation
